---
title: "DEA 1 - Budweiser"
author: "Shijo Joseph & Douglas Yip"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Budweiser

## Executive Summary

The purpose of this review is to provide you white space options based on the category insights to 558 Breweries across the US.  
Our insights that we are sharing are based on the ABV and IBU content by State.
Our review recommends that the following  opportuntities that your company should consider
  1) XXX
  2) YYY
  3) ZZZ

```{r load packages setup, include=FALSE}
#load different libraries and function to enable us to complete this analysis
library(readr)
library(tidyr) #enables piping
library(dplyr)#enables enhancement to cleanse data
library(ggplot2)
library(caret) #enables classification and regression training
library(e1071)
library(class)
library(stringr)



```

```{r load data sets}
#uploads the data into R to start analysis from Github

Breweries <- read.csv("https://raw.githubusercontent.com/sajoseph-smu/Case-Study-1---Budwieser/main/Breweries.csv", header = T) 

summary(Breweries)
Beers <- read.csv("https://raw.githubusercontent.com/sajoseph-smu/Case-Study-1---Budwieser/main/Beers.csv", header = T)
summary(Beers)

```



### 1. How many breweries are present in each state?
```{r }
#This code will identify the number of breweries for each state
#How many Breweries are in each state?

SummaryBrewery <- Breweries %>% count(State, sort = TRUE)
names(SummaryBrewery)[2] <- "Brewery_Count"
#Summary of Breweries 
SummaryBrewery

# Top 25 States with Most Breweries


top_n(SummaryBrewery,25,Brewery_Count) %>% ggplot(mapping = aes(x = reorder(State,-Brewery_Count), y = Brewery_Count, fill = State)) + geom_bar(stat = 'identity') + ggtitle("Top 25 States by Brewery Count") + xlab("States") + ylab("Brewery Count") +geom_text(aes(label = format(Brewery_Count,big.mark =',')), position = position_dodge(width = .9), vjust = -.25)


# Bottom 25 States with Most Breweries


top_n(SummaryBrewery,25,-Brewery_Count) %>% ggplot(mapping = aes(x = reorder(State,Brewery_Count), y = Brewery_Count, fill = State)) + geom_bar(stat = 'identity') + ggtitle("Bottom 25 States by Brewery Count") + xlab("States") + ylab("Brewery Count") +geom_text(aes(label = format(Brewery_Count,big.mark =',')), position = position_dodge(width = .9), vjust = -.25)


```

### 2.   Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  
```{r }
#This code will merge the Data with assigning each beer with the brewery and location

#Merge Data
#outer join (keep all observations from both dataframes)
names(Beers)[5] <- "Brew_ID"
names(Breweries)[2] <- "Brewery_Name"
bbdf <- merge(Beers,Breweries, by = "Brew_ID")

#Head
bbdf %>% head(6)

#tail
bbdf %>% tail(6)
```


### 3.   Address the missing values in each column.
``` {r}
#Cleanse data
#We chose to omit the NA records as a method to clean the data
bbdf_clean = na.omit(bbdf)
```


### 4. Median of alcohol content and bitterness by each state.
```{r }
#This code will create a table of Medians for ABV and IBU for each state and graphically display a bar chart 


# created a clean table with stats for ABV and IBU
bbdf_clean_stats = bbdf_clean %>% group_by(State) %>% summarise(Mean_ABV=mean(ABV), Max_ABV=max(ABV), Min_ABV=min(ABV), Median_ABV=median(ABV), Std_ABV=sd(ABV),Mean_IBU=mean(IBU), Max_IBU=max(IBU), Min_IBU=min(IBU), Median_IBU=median(IBU), Std_IBU=sd(IBU))

# Clean table with Medians for ABV and IBU
bbdf_clean_Medians = bbdf_clean_stats %>% select(State, Median_ABV, Median_IBU)
bbdf_clean_Medians


# Top 25 States by Median ABV
top_n(bbdf_clean_Medians,25,Median_ABV) %>% ggplot(mapping = aes(x = reorder(State,Median_ABV), y = Median_ABV, fill = State)) + geom_bar(stat = 'identity') + ggtitle("Top 25 States by Median ABV") + xlab("States") + ylab("ABV") +ylim(0,.08) +geom_text(aes(label = format(Median_ABV,big.mark =',')), position = position_dodge(width = .6), hjust = -.25) + coord_flip()

# Bottom 25 States by Median ABV
top_n(bbdf_clean_Medians,25,-Median_ABV) %>% ggplot(mapping = aes(x = reorder(State,-Median_ABV), y = Median_ABV, fill = State)) + geom_bar(stat = 'identity') + ggtitle("Bottom 25 States by Median ABV") + xlab("States") + ylab("ABV") +ylim(0,.08) +geom_text(aes(label = format(Median_ABV,big.mark =',')), position = position_dodge(width = .6), hjust = -.25) + coord_flip()

# Top 25 States by Median IBU
top_n(bbdf_clean_Medians,25,Median_IBU) %>%  ggplot(mapping = aes(x = reorder(State,Median_IBU), y = Median_IBU, fill = State)) + geom_bar(stat = 'identity') + ggtitle("Top 25 States by Median IBU") + xlab("States") + ylab("IBU") +ylim(0,70) +geom_text(aes(label = format(Median_IBU,big.mark =',')), position = position_dodge(width = .6), hjust = -.25) + coord_flip()

# Bottom 25 States by Median IBU
top_n(bbdf_clean_Medians,25,-Median_IBU) %>% ggplot(mapping = aes(x = reorder(State,-Median_IBU), y = Median_IBU, fill = State)) + geom_bar(stat = 'identity') + ggtitle("Bottom 25 States by Median IBU") + xlab("States") + ylab("IBU") +ylim(0,70) +geom_text(aes(label = format(Median_IBU,big.mark =',')), position = position_dodge(width = .6), hjust = -.25) + coord_flip()

```

### 5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r }
#This code will identify which State has the highest ABV and IBU beers.


# State with the max ABV
bbdf_clean %>% filter(ABV == max(ABV))

#KY has the Max ABV of .125 based on London Balling from the Against the Grain Brewery

#State with the max IBU
bbdf_clean %>% filter(IBU == max(IBU))


#OR has the Max IBU of 138 based on the Bitter Bitch Imperial IPA from the Astoria Brewing Company
```


### 6.Comment on the summary statistics and distribution of the ABV variable.
```{r }
#This code will create a summary statistic of the ABV variable

# already have that in the bbdf_clean_stats 

# created a clean table with stats for ABV 
bbdf_clean_stats = bbdf_clean %>% group_by(State) %>% summarise(Mean_ABV=mean(ABV), Max_ABV=max(ABV), Min_ABV=min(ABV), Median_ABV=median(ABV), Std_ABV=sd(ABV))


```

### 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer
```{r }

# Scatter Plot to show relationship for the ABV and IBU

bbdf_clean %>% ggplot(mapping = aes(x = ABV, y = IBU)) + geom_point(mapping = aes(x = ABV, y = IBU, color = State), position = "jitter") +  geom_smooth(method='lm', formula= y~x) + ggtitle("ABV vs IBU") + xlab("ABV") + ylab("IBU") 

# there is a linear positive correlation between ABV and IBU based on the linear model of the data set.



```



### 8.  Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually.
```{r  }



# building the Ales and IP data frame and creating a flag for OtherAles and IPA 

bbdf_Ales = bbdf_clean %>% filter(str_detect(Style,"(IPA| Ale)"))

bbdf_Ales = bbdf_Ales %>% mutate(IPAFlag = case_when(str_detect(bbdf_Ales$Style,"IPA") == TRUE ~ "IPA", TRUE ~ "Other Ales" ))

bbdf_Ales$IPAFlag = as.factor(bbdf_Ales$IPAFlag)

# find the best K 
set.seed(1)
iterations = 100
numks = 15
splitPerc = .70

masterAcc = matrix(nrow = iterations, ncol = numks)

for(j in 1:iterations)
{
  trainIndices = sample(1:dim(bbdf_Ales)[1],round(splitPerc * dim(bbdf_Ales)[1]))
  train = bbdf_Ales[trainIndices,]
  test = bbdf_Ales[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(4,5)],test[,c(4,5)],train$IPAFlag, prob = TRUE, k = i)
    CM = confusionMatrix(table(classifications,test$IPAFlag))
    masterAcc[j,i] = CM$overall[1]
  }
  
}

MeanAcc = colMeans(masterAcc)


# find out the best K and accuracy  and plot it
which.max(MeanAcc)
max(MeanAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")
abline(v = 5, col = 'green')
abline(h = .8589, col = 'green')

     
# run the knn with the best K
# K = 5

    classifications = knn(train[,c(4,5)],test[,c(4,5)],train$IPAFlag, prob = TRUE, k = 5)
    CM = confusionMatrix(table(classifications,test$IPAFlag))
    CM
  

# Douglas.. can you write up the conclusion here :) 



```

### 9..  Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 
```{r  }

# creating a map plot of the counts of IPA's and Other Ales 
library(ggplot2)
library(maps)
library(dplyr)
library(mapproj)
#install.packages("tidycensus")
library(tidycensus)


## Run the below code if your are changing the census key
##census_api_key("60e6bde9f7a79e136e60fbbb59ce47e85938deb5", install = TRUE)
##readRenviron("~/.Renviron")






lookup = data.frame(abb = state.abb, State = state.name)
bbdf_map = merge(bbdf, lookup, by.x = "State", by.y = "abb")
bbdf_map
colnames(bbdf_map)[1] = "StateCode"
colnames(bbdf_map)[11] = "State"

bbdf_map$region <- tolower(bbdf_map$State)
bbdf_map_summary = bbdf_map %>% group_by(region) %>% summarize(cnt = n(), meanABV = mean(ABV, na.rm = TRUE), meanIBU = mean(IBU,  na.rm = TRUE))


# Getting population data from the ACS for 5 year ( latest year is 2019)
v19 <- load_variables(2019, "acs5")
pop2019 = get_acs(geography = "state", variable = "B01003_001", year = 2019)
 
# creating a region and a per 100K count for population estimate
pop2019$region = tolower(pop2019$NAME)
pop2019$per100k = pop2019$estimate / 100000


bbdf_map_summary2 = bbdf_map_summary %>% merge(pop2019[, c("region","per100k")], by="region", all.x=T)

bbdf_map_summary2$perCapita = bbdf_map_summary2$cnt / bbdf_map_summary2$per100k 

states <- map_data("state")
map.df <- bbdf_map_summary2 %>% merge(states, by="region", all.x=T)

map.df <- map.df[order(map.df$order),]

# below is the code for creating the Mean ABV by State
# ------

ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill= meanABV))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle("Mean ABV by State")+
  coord_map()
# ------

# below is the code for creating the Other Ales heat map

# ----

ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=meanIBU))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle(" Mean IBU by State")+
  coord_map()

#--- 



# below is the code for creating the Beer Count heat map

# ----

ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=cnt))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle("Beer Count by State")+
  coord_map()

# 


# below is the code for creating the Beer Count heat map

# ----

ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=perCapita))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle("Per Capita by State")+
  coord_map()

# 




s



```